\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-19 arch.tex
updates to arch with embedded comments for FP

\subsubsection{Synchronization identify internal state evolution}
\label{sec:arch:synch}

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal. Even in this
condition, \rx requires each reactor to produce at every tick its {\em
  internal} state. For those reactors not owning {\em internal}
timelines, \rx makes sure that {\em external} observations are taken
into account for every tick as shown in \cite{py10}. Such a
requirement is enforced to ensure that all reactors share a consistent
view of the world with advancing time.

{\em External} state updates are managed by introducing these
observation in the plan structure as facts and ensure that they remain
compatible with the current plan mainainted by the reactor. The
challenging part of synchronization is to decide the {\em internal}
state values for the current tick. In the Shopping example for example
we have the following code snippet which helps identify when we can
\texttt{Buy} a product:

@@

If \texttt{object.location} is {\em external} and the \texttt{Agent}
is {\em internal}, a robot may encounter the situation where it is at
the location where it can buy one (possibly among others)
product. This results om a decision problem related to the state value
of the \texttt{Agent} timeline where the robot can either {\em Buy} a
specific product now, wait a little longer before doing so or just
simply not buy any product at this location.

In the \eu based reactor this decision requirement is a new type of
flaw which forces full identification of internal state of the reactor
for the current tick. Using this new flaw we can describe the
synchronization process as:

@@

\item Propagate this information in the plan database following the
  model $\mathcal{M}$ of this reactor.

\item Resolve the current state value of each internal timelines.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{\small Illustration of the synchronization flaws in a reactor. The
    reactor receive new observations when they are produced by the
    owner(s) of its internal timelines. The line after the last token
    of each timeline represent the domain of possible values for the
    end of this token. At every tick $\tau$ the reactor needs to
    integrate the {\em External} state information it received and --
    based on its model $\mathcal{M}$ -- resolve its {\em Internal}
    state that will then be provided by the architecture to other
    reactors using these state variables.}
  \label{fig:synch:flaw}
\end{figure}

We introduce this requirement as a new type of flaw for the \eu
framework defined as follow:

\begin{definition}
  \label{def:csf}
  A {\em current state flaw} indicates that a reactor {\em internal}
  timeline does not provide a fully defined state value for the
  execution frontier $\tau$. Such a flaw can be result of either of
  the following reasons:

  \begin{itemize}
  \item The absence of a token in this timeline which overlaps the
    instant $\tau$.
  \item An open decision between terminating the last token for this
    timeline or pushing its end in the future.
  \end{itemize}

The core idea between this flaw is to ensure that at any single tick
the {\em synchronization} solver will identify all the {\em internal}
timelines for which the current state value is yet to be
identified. We handle their resolution based on the following
proposition.

\begin{proposition}
  \label{prop:csf:resolve}
  A {\em current state flaw} can be resolved using one of the
  following choices which are evaluated in the following order:

  \item Extend the previous state value to end after the execution 
    frontier  (\ie restrict its end time to $[\tau+1, \infty)$). %\comment{first use of
    % $\tau$, so needs to be defined earlier}

  \item Create and insert a new token in this timeline that will start
    at the current tick $\tau$ (attempt this for each possible token
    type for this timeline if necessary).

\end{enumerate}
\end{proposition}
Intuitively the first choice enforces an intertial value assumption
where the reactor will tend to maintain its current as long as the
model and external observations allow it. The second choice favorise
the advance of the reactor current plan if any; this direct the {\em
  synchronization} decision toward assuming that its current plan is
properly executed. Finally if none of these two choices did provide a
valid solution -- meaning that both lead to an inconsistant plan --
the reactor tries all the possible state values remaining for this
timeline until it finds one that allows it to identify its state as
required by \rx architecture.

This flaw forces the reactor to identify its internal state
for the current tick. By using this flaw one can describe the
synchronization process as:

\begin{enumerate}
\item Integrate the external state as provided by the owner of each
  external timeline into the plan database.
\item Propagate this information in the plan database following the
  model $\mathcal{M}$ of this reactor.
\item Resolve the current state value of each internal timelines based
  on Proposition \ref{prop:csf:resolve}.
\end{enumerate}

@@

Steps $2$ \& $3$ are handled by the \eu solver from algorithm
\ref{alg:europa:solve}. We only enforce that the {\em current state
  flaws} are selected last by the algorithm in order to follow the
sequencing given above. As {\em synchronization} focusses on the
execution frontier wee also filter out all flaws that are outside of
the current tick scope. This mean that if this flaw relates to a token
that is {\em necessarily} ending before $\tau$ or starting after
$\tau+1$ then the solver will ignore it. This helps to reduce the
scope of the synchronization to exclusively the current state of the
reactor and reduce dramatically the search space for the {\em
  synchronization} solver. In order to do so we need to assume 
that the current state value of an {\em internal} timeline does not 
depend on the future or more accurately that any choices made during 
this {\em synchronization} will not lead to a future inconsistency. 
Such an assumption impacts the set of possible domains the reactor 
can support while remaining complete. Take for  example this rule from 
our \texttt{Shopping} model :

@@

While this model is perfectly acceptable for planning it becomes
problematic when execution is taken into account especially since it
puts a strong emphasis between an evolving plan and an unknown future
(\texttt{Going} to a location and the {\em expected} future outcome of
ending at this location, ties it strongly to the constraint in line
$5$ above).  So for example, assume that the \texttt{Agent.location}
is {\em external} to the reactor which receives the observation
\texttt{Going(Home,SuperMarket)} which synchronization has resolved by
producing the observation \texttt{Go(Home, SuperMarket)} in the
\texttt{Agent} timeline. This would imply that the only possible
outcome on \texttt{Agent.location} is to be \texttt{At(SuperMarket)}
in the directly foreseeable future. In the real world the fact that we
are \texttt{Going} to a certain location does not necessarly imply
that we will succeed on ending up there. In one extreme case due to an
exogenous event, the road to the \texttt{SuperMarket} is closed which
could result in an alternate observation that the robot is
\texttt{At(7th st. \& Main)}. This observation contradicts the
existing plan the reactor is maintaining. Further, it is important to
understand the temporal direction of the \emph{causal link} that is
triggering the inconsistency. At the time the robot observes this
contradicting the \texttt{At} predicate, the rule that generated this
conflict is connected to past observations \texttt{Going} and
\texttt{Go} which cannot be easily retracted without a potential
global impact to the plan implying a huge cost in back-propagation
that would be problematic.

More generally tying a token to future outcomes in a model that is
meant to be confronted by execution is problematic. It might be valid
if one can guarantee that an outcome is predisposed to occur (for
example the reactor managing \texttt{Agent.location} timeline
guarantees that the outcome of the \texttt{Going} will {\em always}
result on ending at the traget location). However, given the very
nature of the planning problem and the ability to search outcomes and
replan, the model should be written so that a future projection can be
manipulated. %  while keeping in mind that predicting the future is not
% absolute. 
In this case one could alter this part of the model in order to give
the reactor more flexibility as follows:

@@

We do so by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ which ensures that the last \texttt{Going} ends up in the
location the agent actually wants to be at. This first level of
relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first one fail to reach the
objective.  % The problem can even be relaxed further if needed but
% this relaxation gives already more flexibility for the reactor to
% recover from local failures.
As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions. Since such a process
is required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be carefully designed.

One critical aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. And by doing so we allow for graceful degradation
where other reactors can readapt their plans as they receive this
observation.

Finally, at the engine level careful choices were made in 
both the focus of the search -- limited only to tokens that 
can potentially overlap the current tick -- and heuristics where 
the {\em current state flaw} is considered last during the search
process. The later limits the amount of backtracking for a solution 
-- or at least how deep one needs to backtrack in a decision tree. 
Similarly the sequence in which we evaluate the possible resolution 
for such a flaw -- presented in Proposition \ref{prop:csf:resolve} -- 
was selected according to  some simple assumptions that any external 
observation that is compatible with an existing token in the current
plan reflects the execution of the plan itself. In other words this
means that our solver always prefer to use tokens already existing in
the plan in order to identify its current internal state. By doing so
we directs the synchronization search by exploiting the structure of
the plan produced during previous deliberation steps. This implies
that a deliberation steps have been done and did impact the plan
structure and calls for a description on how our \eu reactor resolve
such steps which we are now going to describe further.

>>>>

\subsubsection{Synchronization and Identifying Internal State
  Evolution}
\label{sec:arch:synch}

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. In
doing so we assume that we have a reactor that has no future
goal. Under such a condition, \rx requires each reactor to produce at
every tick its {\em internal} state. For those reactors not owning
{\em internal} timelines, \rx makes sure that {\em external}
observations are taken into account for every tick as shown in
\cite{py10}. Such a requirement is enforced to ensure that all
reactors share a consistent view of the world with advancing time.

{\em External} state updates are managed by introducing these
observations in the plan structure as facts and ensuring that they
remain compatible with the current plan mainainted by the reactor. The
challenging part of synchronization is to decide the {\em internal}
state values for the current tick. In the Shopping example for example,
we have the following code snippet which helps identify when we can
\texttt{Buy} a product:

@@

If \texttt{object.location} is {\em external} and \texttt{Agent} is an
{\em internal} timeline, a robot may encounter the situation where it
is at the location where it can buy one (possibly among others)
product. This results in a decision-making problem related to the
state value of the \texttt{Agent} timeline; the robot can either {\em
  Buy} a specific product now, wait a little longer before doing so or
just simply not buy any product at this location.

In \texttt{T-REX}'s \eu reactor this decision-making requirement is a
new type of flaw which forces identification \kcomment{identification
  of what? Probably need a little more explanation here.} of internal
state of the reactor for the current tick. This flaw can be defined
as:

\begin{definition}
  \label{def:csf}
  A {\em current state flaw} indicates that a reactor's {\em internal}
  timeline does not provide a fully defined state value for the
  execution frontier $\tau$. Such a flaw can be the result of either:

  \begin{itemize}
  \item The absence of a token in this timeline which overlaps the
    execution frontier $\tau$.
  \item An open decision between terminating the last token for this
    timeline or extending its end in the future.
  \end{itemize}

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{\small Illustration of synchronization flaws in a
    reactor. The reactor receives new observations when they are
    produced by the owner(s) of its internal timelines. The line
    \kcomment{what line???} after the last token of each timeline
    represent the domain of possible values for the end of this
    token. At every tick $\tau$ the reactor needs to integrate the
    {\em External} state information it received and -- based on its
    model $\mathcal{M}$ -- resolve its {\em Internal} state that will
    then be provided by the agent to other reactors using these state
    variables.}
  \label{fig:synch:flaw}
\end{figure}

The key idea is to ensure that at any single tick the {\em
  synchronization} solver will identify all the {\em internal}
timelines for which the current state value is yet to be
identified. We handle their resolution based on the following
proposition:

\begin{proposition}
  \label{prop:csf:resolve}
  A {\em current state flaw} can be resolved using one of the
  following choices which are evaluated in order:

  \item Extend the previous state value to end after the execution
    frontier (\ie restrict its end time to $[\tau+1, \infty)$).

  \item Create and insert a new token in this timeline that will start
    at the current tick $\tau$ and attempt this for each possible
    token type for this timeline if necessary.

\end{enumerate}
\end{proposition}

Intuitively the first choice enforces an \emph{inertial value
  assumption} where the reactor will tend to maintain its current
state as long as the model and external observations allow for it. The
second choice favors the advance of the reactor's current plan if any;
this directs the {\em synchronization} decision toward assuming that
its current plan is properly executed. Finally if none of these two
choices provides a valid solution -- \ie lead to an inconsistant state
-- the reactor tries all the possible state values remaining for the
timeline until it finds one that brings plan consistency.

% allows it to identify its state as required by \rx architecture.

% The flaw forces the reactor to identify its internal state for the
% current tick. By using this flaw one can describe the synchronization
% process as:

The synchronization process in turn can be described as:

@@

\item Propagate this information in the plan database using model
  $\mathcal{M}$ of this reactor.

\item Resolve the current state value of each internal timelines based
  on Proposition \ref{prop:csf:resolve}.

@@

% \begin{enumerate}
% \item Integrate the external state as provided by the owner of each
%   external timeline into the plan database.
% \item Propagate this information in the plan database following the
%   model $\mathcal{M}$ of this reactor.
% \end{enumerate}

@@

Steps $2$ \& $3$ above are handled by the \eu solver from algorithm
\ref{alg:europa:solve} which ensures that the {\em current state
  flaws} are selected last so as to follow the sequence given
above.

As {\em synchronization} focuses on the execution frontier it also
filters out all flaws that are outside of the current tick's scope.
If the flaw correlates to a token that is {\em necessarily} ending
before $\tau$ or starting after $\tau+1$ then the solver ignores
it. Such a strategy helps to reduce the scope of synchronization
exclusively to the current state of the reactor and in turn reduces
the search space for the {\em synchronization} solver.  To do so we
need to assume that the current state value of an {\em internal}
timeline does not depend on the future or more accurately that any
choice made during this {\em synchronization} will not lead to a
future inconsistency.  Such an assumption impacts the set of possible
domains the reactor can support while remaining complete. Take for
example this rule from our \texttt{Shopping} model:

@@

While this model is perfectly acceptable for planning it becomes
problematic when execution is taken into account especially since it
puts a strong emphasis between an evolving plan and an unknown future;
\texttt{Going} to a location and the {\em expected} future outcome of
ending at this location, ties it strongly to the constraint in line
$5$ above. For instance, assume that the \texttt{Agent.location} is
{\em external} to the reactor which receives the observation
\texttt{Going(Home,SuperMarket)} which synchronization has resolved by
producing the observation \texttt{Go(Home, SuperMarket)} in the
\texttt{Agent} timeline. This would imply that the only possible
outcome on the \texttt{Agent.location} timeline, is to be
\texttt{At(SuperMarket)} in the directly foreseeable future.

In the real world the fact that the robot is \texttt{Going} to a
certain location does not necessarly imply that it will succeed in
ending up there. Due to an exogenous event, for example, the road to
the \texttt{SuperMarket} could be closed, which could result in an
alternate observation that the robot is \texttt{At(7th st. \& Main)}
contradicting the existing plan. Further, it is important to
understand the temporal direction of the \emph{causal link} that is
triggering the inconsistency. At the time the robot observes the
contradiction of the \texttt{At} predicate, the rule that generated
this conflict is connected to past observations \texttt{Going} and
\texttt{Go}. These cannot be retracted easily without a potential
impact to the plan implying in terms of a cost in back-propagation.

More generally tying a token to future outcomes in a model that is
meant to be confronted by execution is problematic. It might be valid
if one can guarantee that an outcome is predisposed to occur (for
example the reactor managing the \texttt{Agent.location} timeline
guarantees that the outcome of the \texttt{Going} will {\em always}
result on ending at the traget location). However, given the very
nature of the planning problem and the ability to search outcomes and
to replan, the model should be written so that a future projection can
be manipulated.

One could therefore alter this part of the model to give the reactor
more flexibility as follows:

@@

We do so by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ which ensures that the last \texttt{Going} ends up in the
location the agent actually wants the robot to be at. This first level
of relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first one fail to reach the
objective.

As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions \kcomment{has a unit
  decision been explained in Europa?}. Since such a process is
required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be designed carefully.

One critical aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the \rx agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. By so doing we allow for graceful degradation where
other reactors can readapt their plans on receipt of this observation.

Finally, at the engine level careful choices were made in both the
focus of the search -- limited only to tokens that can potentially
overlap the current tick -- and heuristics where the {\em current
  state flaw} is considered last during the search process. The later
limits the amount of backtracking for a solution -- or at least how
deep one needs to backtrack in a decision tree.  Similarly the
sequence in which we evaluate the possible resolution for such a flaw
-- presented in Proposition \ref{prop:csf:resolve} -- was selected
according to some simple assumptions, that any external observation
that is compatible with an existing token in the current plan reflects
the execution of the plan itself. In other words this means that our
solver always prefers to use tokens \emph{already existing} in the
plan as a way to identify current internal state. By doing so
synchronization exploits the structure of the plan produced during
previous deliberation steps.

% This implies that a deliberation steps
% have been done and did impact the plan structure and calls for a
% description on how our \eu reactor resolve such steps which we are now
% going to describe further.
